Evolution of the AGN UV luminosity function from redshift 7.5

### 2018-03-21 Paper/paper.tex

\begin{table*}
  % In published version of this table, we could remove file names and add data homogenisation in Notes.
  % Add z ~ 7 qso when you add it to analysis
  % See data.tex for a version of this table that includes file names
  % Should we mark Giallongo, Mortlock, and Banados entries to say that these are sometimes excluded from analysis?
  % Remove notes?
  \caption{Quasar data sets}
  \label{tab:samples}
  \begin{tabular}{crcllrSl}
    \hline
    & ID & $z$ & Survey & Reference & Number & {Area} & Notes \\
    & & & & & of quasars & {(deg$^2$)} & \\
    \hline
    1. & 13 & 0.1--2.2 & SDSS DR7 & \citet{2006AJ....131.2766R} & 48664 & 6248.0 & \\
    2. & 15 & 0.4--2.6 & 2SLAQ SGP & \citet{2009MNRAS.392...19C} & 3663 & 64.2 & \\
    3. & 15  & 0.4--2.6 & --- NGP & \citet{2009MNRAS.392...19C} & 8153 & 127.7 & \\
    4. &  1 & 2.2--3.5 & BOSS DR9 & \citet{2013ApJ...773...14R} & 23301 & 2236.0 & \\
    5. & 13 & 3.7--4.7 & SDSS DR7 & \citet{2006AJ....131.2766R} & 1785 & 6248.0 & Restricted to $z<4.7$ \\
    6. & 17 & 4.7--5.4 & SDSS+WISE & \citet{2016ApJ...829...33Y} & 99 & 14555.0 & Overlaps with 8 for $M_{1450}<-26.73$ \\
    7. &  8 & 4.7--5.1 & SDSS DR7 & \citet{2013ApJ...768..105M} & 148 & 6248.0 & \\
    8. &  8 & 4.7--5.1 & --- Stripe 82 & \citet{2013ApJ...768..105M} & 52 & 235.0 & \\
    9. &  8 & 5.1--5.5 & --- DR7 Extended & \citet{2013ApJ...768..105M} & 28 & 6248.0 & \\
    10. & 8 & 5.1--5.4 & ---  Stripe 82 Extended & \citet{2013ApJ...768..105M} & 10 & 235.0 & \\
    11. & 6 & 3.7--4.8 & NDWFS & \citet{2011ApJ...728L..26G} & 12 & 1.71 & \\
    12. & 6 & 3.8--5.1 & DLS & \citet{2011ApJ...728L..26G} & 12 & 2.05 & \\
    13. & 7 & 4.1--6.3 & CANDELS GOODS-S & \citet{2015AA...578A..83G} & 19 & 0.047 & Rescaled selection probabilities \\
    14. & 18 & 5.8--6.4 & --- Main & \citet{2016ApJ...833..222J} & 24 & 11240.0 & \\
    15. & 18 & 5.9--6.1 & --- Overlap & \citet{2016ApJ...833..222J} & 10 & 4223.0 & \\
    16. & 18 & 5.7--6.1 & --- Stripe 82 & \citet{2016ApJ...833..222J} & 13 & 277.0 & \\
    17. & 10 & 6.0 & CFHQS Deep & \citet{2010AJ....139..906W} & 1 & 4.47 & \\
    18. & 10 & 5.9--6.4 & --- Very Wide & \citet{2010AJ....139..906W} & 16 & 494.0 & \\
    19. & 11 & 6.0--6.2 & Subaru High-$z$ Quasar & \citet{2015ApJ...798...28K} & 2 & 6.5 & \\
    20. & 19 & 6.5--7.4 & UKIDSS & \citet{2011Natur.474..616M} & 1 & 3370.0 & \\
    21. & 19 & 6.5--7.4 & UKIDSS & \citet{2007MNRAS.376L..76V} & 1 & 3370.0 & \\
    22. & 20 & 6.5--7.4 & ALLWISE+UKIDSS+DECaLS & \citet{2018Natur.553..473B} & 1 & 2500.0 & \\
    \hline
  \end{tabular}
\end{table*}

@@

\subsection{Binned luminosity function estimates}

@@

\begin{figure*}
  \begin{center}
    \includegraphics[width=\textwidth,keepaspectratio]{mosaic_small.pdf}
    % mosaic.py 
  \end{center}
  \caption{Luminosity function estimates from $z=0.3$ to $2.6$.  The
    symbols show our inferred binned luminosity functions.  In each
    redshift bin, the black curve shows the best-fit double power law
    model, which is represented by the median of the posterior
    probability distribution function.  Yellow curves show a random
    draw from the posterior probability distribution function.  As
    discussed in the text, luminosity bins that show a decrease in the
    luminosity function at the faint end are deemed to be incomplete
    are are removed from our analysis while fitting the double power
    law model.  These bins are represented by open symbols in this
    figure.  Each panel shows the number of quasars in the respective
    redshift bin.  The number in parenthesis indicates the total
    number of quasars in the bin, before excluding incomplete
    bins.}
  \label{fig:mosaic}
\end{figure*}

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth,keepaspectratio]{rhoqso_withGlobal.pdf}
    % rhoqso.py -- draw_withGlobal()
  \end{center}
  \caption{AGN number density evolution in the global model.  Similar
    to Figure~\ref{fig:evoln_global}, symbols show the best fit
    parameters with 68\% uncertainties from the double power law fits
    to the data in redshift bins.  Black solid curve shows the best
    fit global model, and the coloured curves show a random sample
    drawn from the global posterior. \gk{I still think we should show
      the $-18$ case here.  Mention magnitude limits.  Reduce
      transarency.  Extend plot to at least redshift 7.5.  Increase
      legend handle length to match other figure.}}
  \label{fig:rhoqso}
\end{figure}

\begin{figure}
  \begin{center}
    \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{emissivity.pdf}
    % gammapi.py -- draw_emissivity() 
    \end{tabular}
  \end{center}
  \caption{LyC emissivity of AGN assuming 100\% escape fraction.
    Model luminosity functions are integrated down to $M_{1450}=-18$
    in the left panel and $-20$ in the right panel.  For comparison,
    various other published determinations of the 1~Ry emissivity of
    quasars are also shown in both panels of Figure~\ref{fig:e912}.
    The blue curve in these panels shows the AGN 1~Ry emissivity model
    of \citet{2012ApJ...746..125H}, which is very similar to the model
    of \citet{2007ApJ...654..731H}.  The red symbols with error bars
    show the determination from the fits of
    \citet{2015AA...578A..83G}.  The maroon curve is the published fit
    to the 1~Ry emissivity evolution in the fits of
    \citet{2017MNRAS.466.1160M}.  The green curve is the model of
    \citet{2015ApJ...813L...8M}. \gk{Add uncertainties on these fits,
      perhaps by using emcee.}}
  \label{fig:e912}
\end{figure}

@@

The evolution of the comoving number density of quasars is shown in
Figure~\ref{fig:rhoqso} when the luminosity function is integrated
down to different limits.  Similar to Figure~\ref{fig:evoln_global},
symbols show the best fit parameters with 68\% uncertainties from the
double power law fits to the data in redshift bins.  Black solid curve
shows the best fit global model, and the coloured curves show a random
sample drawn from the global posterior.  Bins with BOSS qsos are not
shown.  The number density of qsos down to the limit of spectroscopic
data is about 1 per $10^5$ cMpc$^3$ at its peak.  This density rapidly
increases at low redshifts and then drops gradually at high redshifts.
Figure~\ref{fig:rhoqso} also shows the familiar downsizing feature in
which the number density of faint qsos peaks are lower redshifts than
that of the bright qsos.  While the number density of qsos with
$M_*<-27$ peaks at $z\sim 2.5$, the number density of qsos with
$M_*<-24$ peaks at $z\sim 2$.  At the faintest luminosity at which
spectroscopic data exist, $M_*<-21$, the number density of qsos peaks
at $z\sim 1$.

@@

The rate of emission of photons of frequency $\nu$ by quasars can be
written as
\begin{equation}
\dot n_\nu = \int dM_\nu \phi(M_\nu) \frac{L_\nu(M_\nu)}{h\nu}.
\end{equation}
Here the integral is over a suitable range of magnitudes, the
monochromatic luminosity $L_\nu(M)$ is related to the (absolute AB)
magnitude by \citep{1983ApJ...266..713O}
\begin{equation}
M_\nu = -2.5\log_{10}L_\nu+51.60,
\end{equation}
and $\phi$ is the luminosity function from Equation~(\ref{eqn:dpl}).
The luminosity at 1~Ry is related to that at 1450~\AA\ by
Equation~(\ref{eqn:sed}) or (\ref{eqn:sed_faint})
\begin{equation}
  L_{912}=L_{1450}\left(\frac{\nu_{912}}{\nu_{1450}}\right)^{-0.61}=L_{1450}\left(\frac{912}{1450}\right)^{0.61},
\end{equation}
depending on the qso luminosity.  The corresponding physical volume
emissivity is 
\begin{equation}
\epsilon_\nu = \dot n_\nu h\nu (1+z)^3.
\label{eqn:epsilon}
\end{equation}
Figure~\ref{fig:e912} shows the resultant 1~Ry emissivity when the
luminosity function is integrated down to $M_{1450}=-20$ (right panel)
and when it is integrated down to $M_{1450}=-18$ (left panel).
Similar to the previous figures, the black open circles in this figure
show the result form individual redshift bins, while the black curve
shown the result from best-fit global model.  Yellow curves show a
random sample from the global posterior distribution.  For comparison,
various other published determinations of the 1~Ry emissivity of
quasars are also shown in both panels of Figure~\ref{fig:e912}.  The
blue curve in these panels shows the AGN 1~Ry emissivity model of
\citet{2012ApJ...746..125H}, which is very similar to the model of
\citet{2007ApJ...654..731H}.  The red symbols with error bars show the
determination from the fits of \citet{2015AA...578A..83G}.  The maroon
curve is the published fit to the 1~Ry emissivity evolution in the
fits of \citet{2017MNRAS.466.1160M}.  The green curve is the model of
\citet{2015ApJ...813L...8M}.  Note that the integration limits for
these other models are not homogeneous.  \citet{2017MNRAS.466.1160M}
integrate the luminosity function down to $M_{1450}=-19$, while
\citet{2015AA...578A..83G} integrate down to $M_{1450}=-18$.
\textbf{[What are the limits for HM12 and MH15?]}

@@

>>>>

\begin{table*}
  % In published version of this table, we could remove file names and add data homogenisation in Notes.
  % Add z ~ 7 qso when you add it to analysis
  % See data.tex for a version of this table that includes file names
  % Should we mark Giallongo, Mortlock, and Banados entries to say that these are sometimes excluded from analysis?
  % Remove notes?
  % Do we really need a sample ID column?
  \caption{Quasar data sets}
  \label{tab:samples}
  \begin{tabular}{crcllrSl}
    \hline
    & ID & $z$ & Survey & Reference & Number & {Area} & Notes \\
    & & & & & of quasars & {(deg$^2$)} & \\
    \hline
    1. & 13 & 0.1--2.2 & SDSS DR7 & \citet{2006AJ....131.2766R} & 48664 & 6248.0 & \\
    2. & 15 & 0.4--2.6 & 2SLAQ SGP & \citet{2009MNRAS.392...19C} & 3663 & 64.2 & \\
    3. & 15  & 0.4--2.6 & --- NGP & \citet{2009MNRAS.392...19C} & 8153 & 127.7 & \\
    4. &  1 & 2.2--3.5 & BOSS DR9 & \citet{2013ApJ...773...14R} & 23301 & 2236.0 & \\
    5. & 13 & 3.7--4.7 & SDSS DR7 & \citet{2006AJ....131.2766R} & 1785 & 6248.0 & Restricted to $z<4.7$ \\
    6. & 17 & 4.7--5.4 & SDSS+WISE & \citet{2016ApJ...829...33Y} & 99 & 14555.0 & Overlaps with 8 for $M_{1450}<-26.73$ \\
    7. &  8 & 4.7--5.1 & SDSS DR7 & \citet{2013ApJ...768..105M} & 148 & 6248.0 & \\
    8. &  8 & 4.7--5.1 & --- Stripe 82 & \citet{2013ApJ...768..105M} & 52 & 235.0 & \\
    9. &  8 & 5.1--5.5 & --- DR7 Extended & \citet{2013ApJ...768..105M} & 28 & 6248.0 & \\
    10. & 8 & 5.1--5.4 & ---  Stripe 82 Extended & \citet{2013ApJ...768..105M} & 10 & 235.0 & \\
    11. & 6 & 3.7--4.8 & NDWFS & \citet{2011ApJ...728L..26G} & 12 & 1.71 & \\
    12. & 6 & 3.8--5.1 & DLS & \citet{2011ApJ...728L..26G} & 12 & 2.05 & \\
    13. & 7 & 4.1--6.3 & CANDELS GOODS-S & \citet{2015AA...578A..83G} & 19 & 0.047 & Rescaled selection probabilities \\
    14. & 18 & 5.8--6.4 & --- Main & \citet{2016ApJ...833..222J} & 24 & 11240.0 & \\
    15. & 18 & 5.9--6.1 & --- Overlap & \citet{2016ApJ...833..222J} & 10 & 4223.0 & \\
    16. & 18 & 5.7--6.1 & --- Stripe 82 & \citet{2016ApJ...833..222J} & 13 & 277.0 & \\
    17. & 10 & 6.0 & CFHQS Deep & \citet{2010AJ....139..906W} & 1 & 4.47 & \\
    18. & 10 & 5.9--6.4 & --- Very Wide & \citet{2010AJ....139..906W} & 16 & 494.0 & \\
    19. & 11 & 6.0--6.2 & Subaru High-$z$ Quasar & \citet{2015ApJ...798...28K} & 2 & 6.5 & \\
    20. & 19 & 6.5--7.4 & UKIDSS & \citet{2011Natur.474..616M} & 1 & 3370.0 & \\
    21. & 19 & 6.5--7.4 & UKIDSS & \citet{2007MNRAS.376L..76V} & 1 & 3370.0 & \\
    22. & 20 & 6.5--7.4 & ALLWISE+UKIDSS+DECaLS & \citet{2018Natur.553..473B} & 1 & 2500.0 & \\
    \hline
  \end{tabular}
\end{table*}

@@

\subsection{Binned luminosity function estimates}
\label{sec:binnedlf}

@@

\begin{figure*}
  \begin{center}
    \includegraphics[width=\textwidth,keepaspectratio]{mosaic_small.pdf}
    % mosaic.py 
  \end{center}
  \caption{Luminosity function estimates in redshift bins from $z=0.1$
    to $6.5$.  The symbols show our inferred binned luminosity
    functions from various data sets: sample ID 13 (red), 15 (green),
    1 (blue), 17 (yellow), 8 (brown), 6 (light blue), 18 (pink), 10
    (orange), and 11 (grey).  See Table~\ref{tab:samples} for
    references.  Open circles in corresponding colours indicate
    excluded incomplete bins for the respective data sets.  In each
    redshift bin, the black curve shows the best-fit double power law
    model, which is represented by the median of the posterior
    probability distribution function.  The grey shaded area shows the
    one-sigma (68.26\%) uncertainty.  Each panel shows the number of
    quasars in the respective redshift bin.  The numbers in
    parentheses indicate the total number of reported quasars in the
    respective redshift bin, before excluding incomplete bins.  See
    Sections~\ref{sec:binnedlf} and \ref{sec:bins} for further
    details.}
  \label{fig:mosaic}
\end{figure*}

@@

\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth,keepaspectratio]{rhoqso_withGlobal.pdf}
    % rhoqso.py -- draw_withGlobal()
  \end{center}
  \caption{AGN number density evolution in the global model.  Similar
    to Figure~\ref{fig:evoln_global}, symbols show the best fit
    parameters with 68\% uncertainties from the double power law fits
    to the data in redshift bins.  Black solid curve shows the best
    fit global model, and the coloured curves show a random sample
    drawn from the global posterior. \gk{I still think we should show
      the $-18$ case here.  Mention magnitude limits.  Reduce
      transarency.  Extend plot to at least redshift 7.5.  Increase
      legend handle length to match other figure. Should we add French
      curve fits to this?}}
  \label{fig:rhoqso}
\end{figure}

\begin{figure*}
  \begin{center}
    \includegraphics[scale=0.65]{emissivity.pdf}
    % gammapi.py -- draw_emissivity() 
  \end{center}
  \caption{The 912~\AA\ emissivity of AGN assuming 100\% escape
    fraction down to a limiting magnitude.  Brown points show the
    emissivity when the best-fit double power law luminosity functions
    at respective redshifts are integrated down to $M_{1450}=-18$.
    Black points show the integration limit when the limit is assumed
    to be $M_{1450}=-21$.  For comparison, various other published
    models of the 1~Ry emissivity of AGN are also shown.  The blue
    curve in these panels shows the AGN 912~\AA\ emissivity model of
    \citet{2012ApJ...746..125H}, which is very similar to the model of
    \citet{2007ApJ...654..731H}.  Red circles with error bars show the
    determination from the fits of \citet{2015AA...578A..83G}.  The
    maroon curve is the published fit to the 1~Ry emissivity evolution
    in the fits of \citet{2017MNRAS.466.1160M}.  The green curve is
    the model of \citet{2015ApJ...813L...8M}.  The grey and brown
    curves show the best-fit form of Equation XX. \gk{Mention MH15 and
      HM12 Lstar limits.  Mention that Akiyama open and filled circles
      are very close.}}
  \label{fig:e912}
\end{figure*}

@@

The evolution of the comoving number density of quasars is shown in
Figure~\ref{fig:rhoqso} when the luminosity function is integrated
down to different limits.  Similar to Figure~\ref{fig:evoln_global},
symbols show the best fit parameters with 68\% uncertainties from the
double power law fits to the data in redshift bins.  Black solid curve
shows the best fit global model, and the coloured curves show a random
sample drawn from the global posterior.  Bins with BOSS qsos are not
shown.  The number density of qsos down to the limit of spectroscopic
data is about 1 per $10^5$ cMpc$^3$ at its peak.  This density rapidly
increases at low redshifts and then drops gradually at high redshifts.
Figure~\ref{fig:rhoqso} also shows the familiar downsizing feature in
which the number density of faint qsos peaks are lower redshifts than
that of the bright qsos.  While the number density of qsos with
$M_*<-27$ peaks at $z\sim 2.5$, the number density of qsos with
$M_*<-24$ peaks at $z\sim 2$.  At the faintest luminosity at which
spectroscopic data exist, $M_*<-21$, the number density of qsos peaks
at $z\sim 1$.  \gk{Discuss comparison of expected number of quasars at
  redshift 7 with UKIDSS.}

@@

The rate of emission of photons of frequency $\nu$ by quasars can be
written as
\begin{equation}
\dot n_\nu = \int dM_\nu \phi(M_\nu) \frac{L_\nu(M_\nu)}{h\nu}.
\end{equation}
Here the integral is over a suitable range of magnitudes, the
monochromatic luminosity $L_\nu(M)$ is related to the (absolute AB)
magnitude by \citep{1983ApJ...266..713O}
\begin{equation}
M_\nu = -2.5\log_{10}L_\nu+51.60,
\end{equation}
and $\phi$ is the luminosity function from Equation~(\ref{eqn:dpl}).
The luminosity at 1~Ry is related to that at 1450~\AA\ by
Equation~(\ref{eqn:sed}) or (\ref{eqn:sed_faint})
\begin{equation}
  L_{912}=L_{1450}\left(\frac{\nu_{912}}{\nu_{1450}}\right)^{-0.61}=L_{1450}\left(\frac{912}{1450}\right)^{0.61},
\end{equation}
depending on the qso luminosity.  The corresponding physical volume
emissivity is 
\begin{equation}
\epsilon_\nu = \dot n_\nu h\nu (1+z)^3.
\label{eqn:epsilon}
\end{equation}
Figure~\ref{fig:e912} shows the resultant 1~Ry emissivity when the
luminosity function is integrated down to $M_{1450}=-20$ (right panel)
and when it is integrated down to $M_{1450}=-18$ (left panel).
Similar to the previous figures, the black open circles in this figure
show the result form individual redshift bins, while the black curve
shown the result from best-fit global model.  Yellow curves show a
random sample from the global posterior distribution.  For comparison,
various other published determinations of the 1~Ry emissivity of
quasars are also shown in both panels of Figure~\ref{fig:e912}.  The
blue curve in these panels shows the AGN 1~Ry emissivity model of
\citet{2012ApJ...746..125H}, which is very similar to the model of
\citet{2007ApJ...654..731H}.  The red symbols with error bars show the
determination from the fits of \citet{2015AA...578A..83G}.  The maroon
curve is the published fit to the 1~Ry emissivity evolution in the
fits of \citet{2017MNRAS.466.1160M}.  The green curve is the model of
\citet{2015ApJ...813L...8M}.  Note that the integration limits for
these other models are not homogeneous.  \citet{2017MNRAS.466.1160M}
integrate the luminosity function down to $M_{1450}=-19$, while
\citet{2015AA...578A..83G} integrate down to $M_{1450}=-18$.  \gk{Add
  discussion about Madau and Khaire's integration limit from Gabor's 8
  March email here.  Also discussion on McGreer's data from Gabor's 12
  March email.}

@@

\gk{Add discussion about Schulze point from Gabor's 8 March email
  here.}
